---
output: pdf_document
---
Relationship between a set of variables and miles per gallon
========================================================

## Summary
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

We are interested in two question: "Is an automatic or manual transmission better for Mpg?" and "Quantifying how different is the Mpg between automatic and manual transmissions?" 
After the data analysis I can say there is a significant relationship between the fuel consumption and the transmission type so the manual transmission is better for Mpg. 
After experimenting with linear models I found that third model describes Mpg variable the best.

### Calculating the mean of each transmission types

```{r}
data(mtcars)
automean<-mean(mtcars$mpg[mtcars$am=="0"])
manualmean<-mean(mtcars$mpg[mtcars$am=="1"])
```
The mean for manual transmission is `r manualmean` Mpg and `r automean` Mpg for automatic transmission. This is a significant difference in the means. Manual transmissions have a higher value so based on mean of Mpg it is better (cheaper) to have manual 
transmission.

### Regression model of transmission type affecting the car's fuel consumption

Let's make a simple linear regression model with Mpg as depending variable and am as explanatory variable. We get following model: 
Mpg = 17.147 + 7.245 * am. Since manual transmission is denoted with am = 1, we can conclude that we can make more Mpg with manual 
transmission.

See the boxplot as fig 1 for visual explanation.
```{r}
fit1 <- lm(mpg~am, data=mtcars)
summary(fit1)
```

### Finding the best model

Let's fit a model with all explanatory variables. R-squared = 0.869
```{r}
fitAll<-lm(mpg~.,mtcars)
summary(fitAll)
```

P values are high so we can use only those variables with low p values, for example p, wt, qsec and am. Actually, we got 0.8579 R-squared for model lm(mpg~hp+wt+qsec+am,mtcars) which is lower than fitAll model, but with a little experimenting, I got the following model with 0.8942 R-squared:
```{r}
fit2<-lm(mpg~hp*wt+qsec+am,mtcars)
summary(fit2)
```

### Comparing models

We got 2 models, second one combining variables describes Mpg the best.

See fig 2 and 3 for visual explanation.
```{r}
anova(fit1, fit2)

```

P value is lower than 0.05 which means that model fit2 is significant improvement over model fit1

## Appendix

### Figure 1
```{r}
m<-transform(mtcars, am=factor(am))
levels(m$am)[1] <- "Automatic"
levels(m$am)[2] <- "Manual"
plot(m$am , m$mpg, main="MPG vs Transmission", xlab="Transmission", ylab="MPG", )
abline(lm(mpg ~ am, m), col="red")
```

### Figure 2
```{r}
layout(matrix(c(1,2,3,4),2,2))
plot(fit1)
```

### Figure 3
```{r}
layout(matrix(c(1,2,3,4),2,2))
plot(fit2)
```


